Figure 2 and figure 2 supplementary material

### 2022-02-09 Figures/Figure2.Rmd
add option to do Fig2B of any TCGA
add dependencies + wrapper_ridgeplot so you can run the code to do teh
ridgeplot in any of TCGA cancer types

```{r}
library(tidyverse)
```

@@

plot_ridgeplot <- function(data, ct){

  # calculate correlations
  data %>% dplyr::select(-c(ParticipantBarcode, Study)) %>% data.matrix() %>% rcorr() -> cor.table

  # select strings to remove from signature names
  substring_remove <- c(paste("_",ct, sep=""),
    "_Tirosh_sigmatrix_SuppFig_3-b",
    "_scRNA-Seq",
    "Fig2ab-",
    "_PBMCs_scRNAseq_sigmatrix",
    "sigmatrix_",
    "_Fig2cd")

@@

ggplot(ct_subset, aes(x = value, y = key, colour = method)) +
  geom_density_ridges2(jittered_points = TRUE, position = "raincloud", 
                       point_size = 0.25, point_alpha = 0.25, alpha = 0.7, rel_min_height = 0.005, point_color = "black") +
  theme_ridges() + ggtitle(ct) +
  annotate("text", x = 1, y = 1:length(cor.ct.sorted), label = paste0("R = ",round(cor.ct.sorted, 2)), colour = "brown") +
  scale_colour_manual(values = cols) +
   theme(axis.text.x = element_text(color = "black", size = 20, angle = 90, hjust = .5, vjust = .5, face = "plain"),
        axis.text.y = element_text(color = "black", size = 20, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
        axis.title.x = element_text(color = "black", size = 12, angle = 0, hjust = .5, vjust = 0, face = "plain"),
        axis.title.y = element_text(color = "black", size = 12, angle = 90, hjust = .5, vjust = .5, face = "plain"),
        plot.title = element_text(color = "black", size = 20)) +
  ylab("") + xlab("% of cell type")

}

@@

read_csv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Julien/gemdecan/data/HE_fractions.csv") %>%
  dplyr::select(-c(...1)) %>%
  mutate_at("ParticipantBarcode", str_remove_all, ".01A*") %>%
  mutate_at("ParticipantBarcode", str_replace_all, "-", ".") -> HE_all_TCGA

@@

signature_dict[["CBSX_LM22"]] <- list(Lymphocytes = c("B.cells.naive", "B.cells.memory",  "T.cells.CD8",
                                                        "T.cells.CD4.naive", "T.cells.CD4.memory.resting",
                                                        "T.cells.CD4.memory.activated", 
                                                        "T.cells.regulatory..Tregs.", "NK.cells.resting",
                                                        "NK.cells.activated"),
                                        Macrophages = c("Monocytes", "Macrophages.M0", "Macrophages.M1", "Macrophages.M2",),
                                        Neutrophils = c("Neutrophils")) # skipped , "Plasma.cells" "T.cells.follicular.helper", "T.cells.gamma.delta" in Lymphos, as well as "Dendritic.cells.resting", "Dendritic.cells.activated" for macros

@@

```{r}
HE_all_TCGA %>% dplyr::filter(Study == "LUAD") %>%  na.exclude() %>%
  arrange(factor(ParticipantBarcode, levels = ParticipantBarcode))  -> HE_LUAD_

Comm_samples <- Reduce(intersect, list(tcga_deconv$Samples,
                                       HE_LUAD_$ParticipantBarcode))
HE_LUAD_ %>% 
  dplyr::filter(ParticipantBarcode %in% Comm_samples) %>% 
  arrange(factor(ParticipantBarcode, levels = Comm_samples)) -> HE_LUAD

tcga_deconv %>%
  rename_at(vars(matches("^CBSX")), ~ paste0("CBSX_", .)) %>% 
  rename_at(vars(matches("__")), ~ str_replace_all(., "__", "_")) %>%
  dplyr::filter(Samples %in% Comm_samples) %>% 
  arrange(factor(Samples, levels = Comm_samples)) -> LUAD_RNA_decon

LUAD_summarized <- ridgeplot_parse(deconv_data = LUAD_RNA_decon, signatures_dictionary = signature_dict)

# Ridgeplot
identical(HE_LUAD$ParticipantBarcode, LUAD_summarized$Samples) %>% stopifnot()
 HE_LUAD %>% bind_cols(LUAD_summarized) %>% select(-c(Samples)) -> LUAD_merged

for (ct in c("Lymphocytes", "Macrophages", "Neutrophils")){
 print(plot_ridgeplot(data = LUAD_merged, ct = ct))
 }
# Suplementary idea: apply the above to all TCGA and add dataset as a plot argument

>>>>

```{r}
library(tidyverse)
library(Hmisc)
library(ggridges)
```

@@

plot_ridgeplot <- function(data){
 data %>% dplyr::select(-c(ParticipantBarcode, Study)) %>% data.matrix() %>% rcorr() -> cor.table
  final_plot <- list()
  for (ct in c("Lymphocytes", "Macrophages", "Neutrophils")){
    # select strings to remove from signature names
    substring_remove <- c(paste("_",ct, sep=""),
      "_Tirosh_sigmatrix_SuppFig_3-b",
      "_scRNA-Seq",
      "Fig2ab-",
      "_PBMCs_scRNAseq_sigmatrix",
      "sigmatrix_",
      "_Fig2cd")

@@

  final_plot[[ct]] <- ggplot(ct_subset, aes(x = value, y = key, colour = method)) +
    geom_density_ridges2(jittered_points = TRUE, position = "raincloud", 
                         point_size = 0.25, point_alpha = 0.25, alpha = 0.7, rel_min_height = 0.005, point_color = "black") +
    theme_ridges() + ggtitle(ct) +
    annotate("text", x = 1, y = 1:length(cor.ct.sorted), label = paste0("R = ",round(cor.ct.sorted, 2)), colour = "brown") +
    scale_colour_manual(values = cols) +
     theme(axis.text.x = element_text(color = "black", size = 12, angle = 90, hjust = .5, vjust = .5, face = "plain"),
          axis.text.y = element_text(color = "black", size = 12, angle = 0, hjust = 1, vjust = 0, face = "plain"),  
          axis.title.x = element_text(color = "black", size = 10, angle = 0, hjust = .5, vjust = 0, face = "plain"),
          axis.title.y = element_text(color = "black", size = 10, angle = 90, hjust = .5, vjust = .5, face = "plain"),
          plot.title = element_text(color = "black", size = 20)) +
    ylab("") + xlab("% of cell type")
  }
  return(final_plot)
}

wrapper_ridgeplot <- function(deconv_data, true_estimates, cancer_abreviation){

  # build the data
  true_estimates %>% dplyr::filter(Study == cancer_abreviation) %>%  na.exclude() %>%
    arrange(factor(ParticipantBarcode, levels = ParticipantBarcode))  -> true_subset_

  Comm_samples <- Reduce(intersect, list(deconv_data$Samples,
                                         true_subset_$ParticipantBarcode))
  true_subset_ %>% 
    dplyr::filter(ParticipantBarcode %in% Comm_samples) %>% 
    arrange(factor(ParticipantBarcode, levels = Comm_samples)) -> true_subset

  deconv_data  %>%
    dplyr::filter(Samples %in% Comm_samples) %>% 
    arrange(factor(Samples, levels = Comm_samples)) -> deconv_subset

  deconv_subset_summary <- ridgeplot_parse(deconv_data = deconv_subset, signatures_dictionary = signature_dict)

  identical(true_subset$ParticipantBarcode, deconv_subset_summary$Samples) %>% stopifnot()
  true_subset %>% bind_cols(deconv_subset_summary) %>% select(-c(Samples)) -> data_toplot

  # calculate correlations
 return(plot_ridgeplot(data_toplot))
}

@@

read_csv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Julien/gemdecan/data/HE_fractions.csv") %>%
  dplyr::select(-c(...1)) %>%
  mutate_at("ParticipantBarcode", str_remove_all, ".01A*") %>%
  mutate_at("ParticipantBarcode", str_replace_all, "-", ".") %>%
  distinct(ParticipantBarcode, .keep_all= TRUE) -> HE_all_TCGA # for the READ samples TCGA.AG.5 & TCGA.AG.3

@@

signature_dict[["CBSX_LM22"]] <- list(Lymphocytes = c("B.cells.naive", "B.cells.memory",  "T.cells.CD8",
                                                        "T.cells.CD4.naive", "T.cells.CD4.memory.resting",
                                                        "T.cells.CD4.memory.activated", 
                                                        "T.cells.regulatory..Tregs.", "NK.cells.resting",
                                                        "NK.cells.activated"),
                                        Macrophages = c("Monocytes", "Macrophages.M0", "Macrophages.M1", "Macrophages.M2"),
                                        Neutrophils = c("Neutrophils")) # skipped , "Plasma.cells" "T.cells.follicular.helper", "T.cells.gamma.delta" in Lymphos, as well as "Dendritic.cells.resting", "Dendritic.cells.activated" for macros

@@

tcga_deconv %>%
  rename_at(vars(matches("^CBSX")), ~ paste0("CBSX_", .)) %>% 
  rename_at(vars(matches("__")), ~ str_replace_all(., "__", "_")) -> tcga_deconv_ridge

ridgeplots_by_cancer <- list()
for (cancer in unique(HE_all_TCGA$Study)){
ridgeplots_by_cancer[[cancer]] <- wrapper_ridgeplot(deconv_data = tcga_deconv_ridge, true_estimates = HE_all_TCGA, cancer_abreviation = cancer)
}

# pdf_name <- paste("test", "ridges", ".pdf", sep = "")
#   ggsave(pdf_name, dpi = "print", width = 210, height = 297, units = "mm", scale = 2)
#
